---
title: "Homework 2"
author: "Tony Lian"
output: html_document
---

```{r setup, include=FALSE}
options(scipen = 10)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
library(moments)
library(lubridate)
library(gridExtra)

# Load data
# Annual returns
ret_ann = read.table("data/returns_annual.txt", skip = 5, header = TRUE) %>%
  rename(date = X.date) %>% data.table()


# Monthly returns
ret_mon = read.table("data/returns_monthly.txt", skip = 4, header = TRUE) %>% 
  data.table()
ret_mon$date = ymd(ret_mon$X.date)

# Daily returns
ret_daily = read.table("data/returns_daily.txt", skip = 5, header = TRUE) %>%
  data.table()
ret_daily$date = ymd(ret_daily$X.date)

```


## 1. Recompute the answer to HW1-6 by simulation

```{r}

rets = list(ret_ann, ret_mon, ret_daily)
periods = c("annual", "monthly", "daily")

ret_stats = as.data.frame(matrix(NA, nrow = 3, ncol = 5))

for (i in 1:3) {
  ret = rets[[i]]
  
  ret_stats[i, 1] = periods[i]
  
  ret_stats[i, 2] = mean(ret$mktret)
  ret_stats[i, 3] = sd(ret$mktret)
  
  ret_stats[i, 4] = mean(ret$bondret)
  ret_stats[i, 5] = sd(ret$bondret)
}

names(ret_stats) = c("period", "mkt_mean", "mkt_sd", "bond_mean", "bond_sd")
ret_stats %>% kable() %>% kable_styling()

```
```{r}
# Simulations
n = 10000

stock_sim = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_sim) = periods

bond_sim = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_sim) = periods

for (i in 1:3) {
  # Stock sim
  stock_sim[, i] = rnorm(n, ret_stats[i, 2], ret_stats[i, 3])
  bond_sim[, i] = rnorm(n, ret_stats[i, 4], ret_stats[i, 5])
  
  p1 = stock_sim %>% ggplot(aes_string(x = names(stock_sim)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'darkred') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 2],
                                           sd = ret_stats[i, 3]),
                  size = 1.5) + 
    ggtitle(paste0("Stock simulation: ", periods[i]))
  
  p2 = bond_sim %>% ggplot(aes_string(x = names(bond_sim)[i])) + 
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'darkblue') + 
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 4], 
                                           sd = ret_stats[i, 5]),
                  size = 1.5) + 
    ggtitle(paste0("Bond simulation: ", periods[i]))
  
  grid.arrange(p1, p2, ncol = 2)        
  
}



data.frame(stock_sim, "", bond_sim) %>% head() %>% kable() %>% kable_styling()


```

We get the same answers as in Assignment 1 since we are calculating shortfall based on the same assumed normal distribution.

## 2. Recompute shortfall using the resampling method

```{r}
# Resample from empirical data
stock_samp = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_samp) = periods

bond_samp = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_samp) = periods

for (i in 1:3) {
  stock_samp[, i] = sample(rets[[i]]$mktret, n, replace = TRUE)
  bond_samp[, i] = sample(rets[[i]]$bondret, n, replace = TRUE)
  
  p1 = stock_samp %>% ggplot(aes_string(x = names(stock_samp)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'pink') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 2],
                                           sd = ret_stats[i, 3]),
                  size = 1.5) + 
    ggtitle(paste0("Stock resampling: ", periods[i]))
  
  p2 = bond_samp %>% ggplot(aes_string(x = names(bond_samp)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'lightblue') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 4],
                                           sd = ret_stats[i, 5]),
                  size = 1.5) + 
    ggtitle(paste0("Bond resampling: ", periods[i]))
  
  grid.arrange(p1, p2, ncol = 2)
  
}

```


```{r}
# Function to calculate shortfall 
shortfall = function(x, lvl) {
  mean(x < lvl)
}

# Function to calculate outperformance
outperf = function(x, lvl) {
  mean(x > lvl)
}

# Calculate Shorfalls
lvls = seq(-.2, .2, .1)


t = data.frame(levels = lvls)

for (i in 1:3) {
  
  # Stocks
  t[, paste0("stock_anl_", periods[i])] = t$levels %>% 
    sapply(function(x) {pnorm(x, ret_stats[i, 2], ret_stats[i, 3])})
  
  t[, paste0("stock_sim_", periods[i])] = t$levels %>%
    sapply(function(x) {shortfall(stock_sim[, i], x)})
  
  t[, paste0("stock_samp_", periods[i])] = t$levels %>%
    sapply(function(x) {shortfall(stock_samp[, i], x)})
  
  
  # Bonds
  t[, paste0("bond_anl_", periods[i])] = t$levels %>% 
    sapply(function(x) {pnorm(x, ret_stats[i, 4], ret_stats[i, 5])})
  
  t[, paste0("bond_sim_", periods[i])] = t$levels %>%
    sapply(function(x) {shortfall(bond_sim[, i], x)}) 
  
  t[, paste0("bond_samp_", periods[i])] = t$levels %>%
    sapply(function(x) {shortfall(bond_samp[, i], x)})
  
  # Plots
  p1 = t %>% ggplot(aes_string(x = 'levels')) + 
    geom_line(aes_string(y = paste0('stock_sim_', periods[i])), 
              size = 1, color = 'darkred') + 
    geom_line(aes_string(y = paste0('stock_samp_', periods[i])),
              size = 1, color = 'pink') +
    ggtitle(paste0("Stock shortfalls: ", periods[i]))
  
  
  p2 = t %>% ggplot(aes_string(x = 'levels')) + 
    geom_line(aes_string(y = paste0('bond_sim_', periods[i])), 
              size = 1, color = 'darkblue') + 
    geom_line(aes_string(y = paste0('bond_samp_', periods[i])),
              size = 1, color = 'lightblue') +
    ggtitle(paste0("Bond shortfalls: ", periods[i]))
  
  grid.arrange(p1, p2, ncol = 2)
  
}


t %>% kable() %>% kable_styling()

```
From empirical observations, we see that using the resampling method, monthly and daily return shortfalls are very comparable, but annual shortfalls differ. Empirical sampling yields much higher shortfall probabilities for negative return levels for stocks and the opposite for bonds. It seems to match our calculations that stocks exhibit negative skewness and bonds exhibit positive skewness.

## 3, 4, 5. Shortfalls - Analytical, Simulation, Resampling

```{r}
# Transform to lognormal

ret_stats_log = as.data.frame(matrix(NA, nrow = 3, ncol = 5))

for (i in 1:3) {
  ret_stats_log[i, 1] = periods[i]
  
  ret_stats_log[i, 2] = mean(log(1 + rets[[i]]$mktret))
  ret_stats_log[i, 3] = sd(log(1 + rets[[i]]$mktret))
  
  ret_stats_log[i, 4] = mean(log(1 + rets[[i]]$bondret))
  ret_stats_log[i, 5] = sd(log(1 + rets[[i]]$bondret))
}

names(ret_stats_log) = c("period", "mkt_mean_log", "mkt_sd_log", "bond_mean_log", "bond_sd_log")
ret_stats_log %>% kable() %>% kable_styling()

```
```{r}
# Simulation and Resampling for 5 periods
t = 5

stock_sim5 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_sim5) = periods

bond_sim5 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_sim5) = periods

stock_samp5 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_samp5) = periods

bond_samp5 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_samp5) = periods

for (i in 1:3) {
  stock_sim_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  stock_samp_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  
  bond_sim_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  bond_samp_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  
  for (j in 1:t) {
    stock_sim_df[, j] = rnorm(n, ret_stats$mkt_mean[i], ret_stats$mkt_sd[i]) + 1
    stock_samp_df[, j] = sample(rets[[i]]$mktret, n, replace = TRUE) + 1
    
    bond_sim_df[, j] = rnorm(n, ret_stats$bond_mean[i], ret_stats$bond_sd[i]) + 1
    bond_samp_df[, j] = sample(log(rets[[i]]$bondret + 1), n, replace = TRUE) + 1
  }
  
  stock_sim5[, i] = stock_sim_df %>% Reduce(`*`, .) - 1
  stock_samp5[, i] = stock_samp_df %>% Reduce(`*`, .) - 1
  
  p1 = stock_sim5 %>% ggplot(aes_string(x = names(stock_sim5)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'darkred') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 2] * t,
                                           sd = ret_stats[i, 3] * sqrt(t)),
                  size = 1.5) + 
    ggtitle(paste0("Stock simulation: ", periods[i]))
  
  p2 = stock_samp5 %>% ggplot(aes_string(x = names(stock_samp5)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'pink') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 2] * t,
                                           sd = ret_stats[i, 3] * sqrt(t)),
                  size = 1.5) + 
    ggtitle(paste0("Stock resampling: ", periods[i]))
  
  
  bond_sim5[, i] = bond_sim_df %>% Reduce(`*`, .) - 1
  bond_samp5[, i] = bond_samp_df %>% Reduce(`*`, .) - 1 
  
  p3 = bond_sim5 %>% ggplot(aes_string(x = names(bond_sim5)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'darkblue') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 4] * t,
                                           sd = ret_stats[i, 5] * sqrt(t)),
                  size = 1.5) + 
    ggtitle(paste0("Stock simulation: ", periods[i]))
  
  p4 = bond_samp5 %>% ggplot(aes_string(x = names(bond_samp5)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'lightblue') +
    stat_function(fun = dnorm, args = list(mean = ret_stats[i, 4] * t,
                                           sd = ret_stats[i, 5] * sqrt(t)),
                  size = 1.5) + 
    ggtitle(paste0("Stock resampling: ", periods[i]))
  
  grid.arrange(p1, p2, p3, p4, nrow = 1)
  
  
}

```

```{r}
t = 5
lvl = 0.2

stock_outperf5 = as.data.frame(matrix(NA, nrow = 3, ncol = 4))
names(stock_outperf5) = c("stock_outperf5", "analytical", "simulation", "resampling")

bond_outperf5 = as.data.frame(matrix(NA, nrow = 3, ncol = 4))
names(bond_outperf5) = c("bond_outperf5", "analytical", "simulation", "resampling")

for (i in 1:3) {
  
  stock_outperf5[i, 1] = periods[i]
  bond_outperf5[i, 1] = periods[i]
  
  #Analytical
  stock_outperf5[i, 2] = pnorm(log(1 + lvl), 
                               ret_stats_log[i, 2] * t, 
                               ret_stats_log[i, 3] * sqrt(t),
                               lower.tail = FALSE)
  
  bond_outperf5[i, 2] = pnorm(log(1 + lvl),
                              ret_stats_log[i, 4] * t,
                              ret_stats_log[i, 5] * sqrt(t),
                              lower.tail = FALSE)
  
  
  
  # Simulation
  stock_outperf5[i, 3] = outperf(stock_sim5[, i], lvl)
  bond_outperf5[i, 3] = outperf(bond_sim5[, i], lvl)
  
  
  # Resampling
  stock_outperf5[i, 4] = outperf(stock_samp5[, i], lvl)
  bond_outperf5[i, 4] = outperf(bond_samp5[, i], lvl)
  
  
}


stock_outperf5 %>% kable() %>% kable_styling()

bond_outperf5 %>% kable() %>% kable_styling()


```

## 6, 7. Relative Shortfall

```{r}
# Simulation and Resampling for 30 periods
t = 30

stock_sim30 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_sim30) = periods

bond_sim30 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_sim30) = periods

stock_samp30 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(stock_samp30) = periods

bond_samp30 = as.data.frame(matrix(NA, nrow = n, ncol = 3))
names(bond_samp30) = periods

for (i in 1:3) {
  stock_sim_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  stock_samp_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  
  bond_sim_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  bond_samp_df = as.data.frame(matrix(NA, nrow = n, ncol = t))
  
  for (j in 1:t) {
    stock_sim_df[, j] = rnorm(n, ret_stats$mkt_mean[i], ret_stats$mkt_sd[i]) + 1
    stock_samp_df[, j] = sample(rets[[i]]$mktret, n, replace = TRUE) + 1
    
    bond_sim_df[, j] = rnorm(n, ret_stats$bond_mean[i], ret_stats$bond_sd[i]) + 1
    bond_samp_df[, j] = sample(log(rets[[i]]$bondret + 1), n, replace = TRUE) + 1
  }
  
  stock_sim30[, i] = stock_sim_df %>% Reduce(`*`, .) - 1
  stock_samp30[, i] = stock_samp_df %>% Reduce(`*`, .) - 1
  
  bond_sim30[, i] = bond_sim_df %>% Reduce(`*`, .) - 1
  bond_samp30[, i] = bond_samp_df %>% Reduce(`*`, .) - 1 
  
}

rel_sim30 = bond_sim30 - stock_sim30
rel_samp30 = bond_samp30 - stock_samp30

for (i in 1:3) {
  p1 = rel_sim30 %>% ggplot(aes_string(x = names(rel_sim30)[i])) +
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'seagreen2') +
    ggtitle(paste0("Relative simulation: ", periods[i]))
  
  p2 = rel_samp30 %>% ggplot(aes_string(x = names(rel_samp30)[i])) + 
    geom_histogram(aes(y = ..density..), bins = 50, fill = 'seagreen4') +
    ggtitle(paste0("Relative resampling: ", periods[i]))
  
  grid.arrange(p1, p2, nrow = 1)
}

```



```{r}
# Relative shortfall

rel_shortfall30 = as.data.frame(matrix(NA, nrow = 3, ncol = 4))
names(rel_shortfall30) = c("Relative shortfall", "analytical", "simulation", "resampling")


for (i in 1:3) {
  
  rel_shortfall30[i, 1] = periods[i]
  
  #Analytical
  rel_shortfall30[i, 2] = pnorm(0, 
                                mean(log(1 + rets[[i]]$bondret) - 
                                       log(1 + rets[[i]]$mktret)) * t, 
                                sd(log(1 + rets[[i]]$bondret) - 
                                     log(1 + rets[[i]]$mktret)) * sqrt(t))
  
  
  
  # Simulation
  rel_shortfall30[i, 3] = shortfall(bond_sim30[, i] - stock_sim30[, i], 0)
  
  # Resampling
  rel_shortfall30[i, 4] = shortfall(bond_samp30[, i] - stock_samp30[, i], 0)
  
  
}


rel_shortfall30 %>% kable() %>% kable_styling()



```


## Barnstable College
```{r}
# Exhibit 1
r = data.frame(T = 0:100)
r$rf = exp(0.06 * r$T)

r$mean_r = exp(.117 * r$T)
#r$sig_up2 = r$mean_r * (1 +  2 * .16 * sqrt(r$T))
#r$sig_dn2 = r$mean_r * (1 - 2 * .16 * sqrt(r$T))
r$sig_up2 = exp(.117 * r$T + 2 * sqrt(r$T))
r$sig_dn2 = exp(.117 * r$T - 2 * sqrt(r$T))



r %>% ggplot() + 
  geom_line(aes(x = T, y = mean_r), color = 'black') + 
  geom_line(aes(x = T, y = rf), color = 'black') + 
  geom_line(aes(x = T, y = sig_up2), color = "seagreen") + 
  geom_line(aes(x = T, y = sig_dn2), color = "seagreen") +
  scale_y_log10()

```

```{r}
# Exhibit 2

r = data.frame(T = 0:100)
r$put = sqrt(r$T/300)
r$shortfall = 1/(r$T + 2)


r %>% ggplot() + 
  geom_line(aes(x = T, y = put), color = "black") +
  geom_line(aes(x = T, y = shortfall), color = "blue")



```