---
title: "Homework 4"
author: "Tony Lian"
output: html_document
---

```{r setup, include=FALSE}
options(scipen = 10)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
library(moments)
library(lubridate)
library(gridExtra)

# Load data
tb = fread("data/TB_7301_1912.csv")
ret = fread("data/STOCK_RETS.csv")
ret$date = as.Date(as.character(ret$date), "%Y%m%d")
ret$tb = tb$t30ret
```

# 1 Estimating E and V by the sample estimates
## a) Compute the sample estimates of E and V.
```{r}
tickers = c("XON", "PG", "PFE", "WMT", "INTC")

stocks = ret %>% select(all_of(tickers))

excess_ret = stocks - ret$tb %*% matrix(1,1,5)

E = colMeans(excess_ret)
print(E)
V = cov(stocks)
print(V)


```

## b) Compute the tangency portfolio weights based on E and V.
```{r}
W = (solve(V) %*% E) / sum(solve(V) %*% E)
print("Weights")

print(t(W))

e_tang = (E %*% W)[1]
print("Expected excess returns")
print(e_tang)

v_tang = (t(W) %*% V %*% W)[1]
print("Variance")
print(v_tang)

sr_tang = e_tang / sqrt(v_tang)
print("Sharpe Ratio")
print(sr_tang)


```
I would not be willing to invest according to these weights. 36% in XON is too high for my taste.


## c) How do the tangency portfolio weights compare to the weights in the minimum variance portfolio?
```{r}
i = matrix(1,5,1)
W0 = (solve(V) %*% i) / (t(i) %*% solve(V) %*% i)[1]

print("Weights")
print(t(W0))

e_mvp = (E %*% W0)[1]
print("Expected excess returns")
print(e_mvp)

v_mvp = (t(W0) %*% V %*% W0)[1]
print("Variance")
print(v_mvp)

sr_mvp = e_mvp/sqrt(v_mvp)
print("Sharpe Ratio")
print(sr_mvp)

```
Lower expected return, and lower Variance, lower SR for the MVP.


## d) Since E is affected by statistical noise. Round E to two decimal places and recompute the tangency portfolio weights using rounded numbers.
```{r}

E_rounded = round(E, 2)

W_rounded = (solve(V) %*% E_rounded) / (t(i) %*% solve(V) %*% E_rounded)[1]

print("Weights")
print(t(W_rounded))

print("Differences")
print(t(W_rounded) - t(W))

```

# 2. Estimating V by the identity matrix
## a) Compute the tangency portfolio weight using the identity matrix I in place of V.
```{r}
I = diag(5)

W = (I %*% E) / (t(i) %*% I %*% E)[1]

print("Weights")
print(t(W))
```
## b) How do these weights change if you round E to two decimal places as before?
```{r}

W_rounded = (I %*% E_rounded) / (t(i) %*% I %*% E_rounded)[1]

print("Weights")
print(t(W_rounded))

print("Differences")
print(t(W_rounded) - t(W))

```

# 3. Estimating E using the CAPM
```{r}
risk_prem = 0.005
betas = matrix(c(0.6, 0.7, 1.2, 0.9, 1.2), nrow = 1)

# CAPM excess returns
E_capm = betas * risk_prem
colnames(E_capm) = tickers

print("CAPM Expected Returns")
print(E_capm)

# Tangency Portfolio Weights
W = (solve(V) %*% t(E_capm)) / sum(solve(V) %*% t(E_capm))

print("Weights")
print(t(W))

```



# 4. Estimating E and V using Bayesian/shrinkage Techniques
```{r}
# Weighted average of E and E_capm, Equal weights
E_b = (E + E_capm) / 2

D = mean(diag(V))[1] * I
V_b = (V + D) / 2

W_b = (solve(V_b) %*% t(E_b)) / (t(i) %*% solve(V_b) %*% t(E_b))[1]

print("Weights")
print(t(W_b))

```

# 5. Dynamic portfolio rebalancing
```{r}

r = ret %>% select(all_of(tickers))
re = r - ret$tb %*% matrix(1, 1, 5)

f1 = which(ret$date >= '1978-01-31' & format(ret$date, '%m') == "01")


colnm = c('ret', 'exret')
rownm = format(ret$date[f1[1]:nrow(ret)], '%Y-%m-%d')
out1 = as.data.frame(matrix(NA, length(rownm), 2, dimnames = list(rownm, colnm)))
out2 = as.data.frame(matrix(NA, length(rownm), 2, dimnames = list(rownm, colnm)))
out3 = as.data.frame(matrix(NA, length(rownm), 2, dimnames = list(rownm, colnm)))
out4 = as.data.frame(matrix(NA, length(rownm), 2, dimnames = list(rownm, colnm)))

for (tt in 1:length(f1)) {
  
  # Subset returns and excess returns
  # Historical
  r1 = r[1:(f1[tt]-1), ]
  re1 = re[1:(f1[tt]-1),]
  
  # Historical estimate
  ER1o = colMeans(re1)
  V1o = cov(r1)
  W1o = solve(V1o) %*% ER1o / sum(solve(V1o) %*% ER1o)
  
  # Capm
  ER2o = t(risk_prem * betas)
  W2o = solve(V1o) %*% ER2o / sum(solve(V1o) %*% ER2o)
  
  # Identity Matrix in place of V
  V3o = diag(1, 5)
  W3o = V3o %*% ER1o / sum(ER1o)
  
  # Bayesian
  avgvar = mean(diag(V1o))
  p = 0.5
  ER4o = p * ER1o + (1-p) * ER2o
  V4o = p * V1o + (1-p) * avgvar * diag(1, 5)
  W4o = solve(V4o) %*% ER4o / sum(solve(V4o) %*% ER4o)
  
  
  # Calculate returns for next year
  r2 = as.matrix(r[f1[tt]:(f1[tt]+11), ])
  re2 = as.matrix(re[f1[tt]:(f1[tt]+11), ])
  out1[(12*tt-11):(12*tt), ] = cbind(r2 %*% W1o, re2 %*% W1o)
  out2[(12*tt-11):(12*tt), ] = cbind(r2 %*% W2o, re2 %*% W2o)
  out3[(12*tt-11):(12*tt), ] = cbind(r2 %*% W3o, re2 %*% W3o)
  out4[(12*tt-11):(12*tt), ] = cbind(r2 %*% W4o, re2 %*% W4o)
  
}
  
we1 = cumprod(out1$ret+1) - 1
we2 = cumprod(out2$ret+1) - 1
we3 = cumprod(out3$ret+1) - 1
we4 = cumprod(out4$ret+1) - 1




avgrets = c(mean(out1$ret), mean(out2$ret), mean(out3$ret), mean(out4$ret))
stdrets = c(sd(out1$ret), sd(out2$ret), sd(out3$ret), sd(out4$ret))
sharpes = c(mean(out1$exret)/sd(out1$ret), mean(out2$exret)/sd(out2$ret), 
            mean(out3$exret)/sd(out3$ret), mean(out4$exret)/sd(out4$ret))


print("Average returns")
print(avgrets)

print("Stddevs")
print(stdrets)

print("Sharpe Ratios")
print(sharpes)

agg_dt = data.table(date = ret$date[61:nrow(ret)],
                    we1 = we1,
                    we2 = we2,
                    we3 = we3,
                    we4 = we4)


agg_dt %>% ggplot(aes(x = date)) + 
  geom_line(aes(y = we1), linetype = 1) +
  geom_line(aes(y = we2), linetype = 2) +
  geom_line(aes(y = we3), linetype = 3) + 
  geom_line(aes(y = we4), linetype = 4)

```