---
title: "Homework 5"
author: "Tony Lian"
output: html_document
---
  
```{r setup, include=FALSE}
options(scipen = 10)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
library(moments)
library(lubridate)
library(gridExtra)
library(TTR)

cols = c("date", "mkt", "bm_h", "bm_l", "ep_h", "ep_l", "cp_h", "cp_l", "yld_h", "yld_l", "yld_z")
aus = read.table("data/australia.txt", skip = 2, header = TRUE)
colnames(aus) = cols

fr = read.table("data/france.txt", skip = 2, header = TRUE)
colnames(fr) = cols

ger = read.table("data/germany.txt", skip = 2, header = TRUE)
colnames(ger) = cols

ita = read.table("data/italy.txt", skip = 2, header = TRUE)
colnames(ita) = cols

jpn = read.table("data/japan.txt", skip = 2, header = TRUE)
colnames(jpn) = cols

uk = read.table("data/uk.txt", skip = 2, header = TRUE)
colnames(uk) = cols

ff = read.table("data/ff_factors_192607_201912.txt", header = TRUE)
colnames(ff) = c('date', 'mkt', 'smb', 'hml', 'rf')

```


# 1. For which countries is the value effect significantly different from zero in the full sample? based on your results, do you agree with the opinion of the casewriter on page 4 of the DFA case?

```{r}
countries = list(aus, fr, ger, ita, jpn, uk)
country_names = c("aus", "fr", "ger", "ita", "jpn", "uk")


hml = matrix(data = NA, nrow = dim(aus)[1], 6)
for (i in 1:length(countries)) {
  
  hml[,i] = countries[[i]]$bm_h - countries[[i]]$bm_l
}

hml = as.data.frame(hml)
colnames(hml) = country_names

# Mean, SD, and T-stat
m.hml = colMeans(hml)
s.hml = apply(hml, 2, sd)
t.hml = sqrt(dim(hml)[1]) * m.hml / s.hml

cbind(rbind("mean", "sd", "t"), rbind(m.hml, s.hml, t.hml) %>% round(4)) %>% 
  kable() %>% kable_styling()

```

# 2. For each country, plot the 5-year moving average the value minus growth returns. Which stocks performed better during the 2007-08 financial crisis, value or growth? Does your answer support the Fama French argument that value premium is due to risk?

```{r}
# Moving average calculations
hml.sma.5 = matrix(data = NA, nrow = dim(hml)[1], dim(hml)[2])

for (i in 1:dim(hml)[2]) {
  # SMA of 5 periods
  hml.sma.5[, i] = SMA(hml[,i], n = 5)

}

hml.sma.5 = as.data.frame(hml.sma.5)
colnames(hml.sma.5) = country_names
hml.sma.5$date = aus$date


p = hml.sma.5 %>% filter(date > '200701' & date < '200901') %>% 
  ggplot(aes(x = as.factor(date))) + 
  geom_bar(aes(y = aus, fill = "aus"), alpha = 0.5, stat = 'identity') +
  geom_bar(aes(y = fr, fill = 'fr'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ger, fill = 'ger'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ita, fill = 'ita'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = jpn, fill = 'jpn'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = uk, fill = 'uk'), alpha = 0.5, stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("HML Returns SMA5")


print(p)

```

Seems like value stocks performed better during this period.

# 3. Augment the international data with HML returns. Estimate the 7x7 correlation matrix of the value-growth spreads. Based on your results do you agree with the opinion of the casewriter of the DFA case?

```{r}
# Subset FF dates

hml$date = aus$date
merged = hml %>% merge(ff[, c('date', 'hml')], by.x = 'date', by.y = 'date') %>%
  rename(us = 'hml')

V1 = cor(merged[,-1])

V1 %>% kable() %>% kable_styling()

```

# 4. Given the correlations, would you recommend for US value investors to go global?

Yes, there seems to be low correlations and so we can improve our Sharpe ratio by diversifying internationally.


# 5. Combine the 7 value-growth spreads into a portfolio with minimum total risk. Compare the portfolio's standard deviation with that of HML. Any surprise?

```{r}
I = matrix(1, 7, 1)
# Weightings of Minimum Variance Portfolio
w1 = solve(V1) %*% I / sum(t(I) %*% solve(V1))

print(t(w1))
```

Very low weighting of 6.5% on US equities. Seems like international stocks offer lower variances.

# 6. Consider a mean-variance investor who currently holds all her $1 million in the US market portfolio. She holds zero position in the risk-free T-bill and she considers such a position optimal, given the historical moments of asset returns. What is the investor's risk aversion $\gamma$? Explain how you obtained your result.

```{r}









```