---
title: "Homework 5"
author: "Tony Lian"
output: html_document
---
  
```{r setup, include=FALSE}
options(scipen = 10)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
library(moments)
library(lubridate)
library(gridExtra)
library(TTR)

cols = c("date", "mkt", "bm_h", "bm_l", "ep_h", "ep_l", "cp_h", "cp_l", "yld_h", "yld_l", "yld_z")
aus = read.table("data/australia.txt", skip = 2, header = TRUE)
colnames(aus) = cols

fr = read.table("data/france.txt", skip = 2, header = TRUE)
colnames(fr) = cols

ger = read.table("data/germany.txt", skip = 2, header = TRUE)
colnames(ger) = cols

ita = read.table("data/italy.txt", skip = 2, header = TRUE)
colnames(ita) = cols

jpn = read.table("data/japan.txt", skip = 2, header = TRUE)
colnames(jpn) = cols

uk = read.table("data/uk.txt", skip = 2, header = TRUE)
colnames(uk) = cols

ff = read.table("data/ff_factors_192607_201912.txt", header = TRUE)
colnames(ff) = c('date', 'mkt', 'smb', 'hml', 'rf')

```


# 1. For which countries is the value effect significantly different from zero in the full sample? based on your results, do you agree with the opinion of the casewriter on page 4 of the DFA case?

```{r}
countries = list(aus, fr, ger, ita, jpn, uk)
country_names = c("aus", "fr", "ger", "ita", "jpn", "uk")


hml = matrix(data = NA, nrow = dim(aus)[1], 6)
for (i in 1:length(countries)) {
  
  hml[,i] = countries[[i]]$bm_h - countries[[i]]$bm_l
}

hml = as.data.frame(hml)
colnames(hml) = country_names

# Mean, SD, and T-stat
m.hml = colMeans(hml)
s.hml = apply(hml, 2, sd)
t.hml = sqrt(dim(hml)[1]) * m.hml / s.hml

cbind(rbind("mean", "sd", "t"), rbind(m.hml, s.hml, t.hml) %>% round(4)) %>% 
  kable() %>% kable_styling()

```

# 2. For each country, plot the 5-year moving average the value minus growth returns. Which stocks performed better during the 2007-08 financial crisis, value or growth? Does your answer support the Fama French argument that value premium is due to risk?

```{r}
# Moving average calculations
hml.sma.5 = matrix(data = NA, nrow = dim(hml)[1], dim(hml)[2])

for (i in 1:dim(hml)[2]) {
  # SMA of 5 periods
  hml.sma.5[, i] = SMA(hml[,i], n = 5)

}

hml.sma.5 = as.data.frame(hml.sma.5)
colnames(hml.sma.5) = country_names
hml.sma.5$date = aus$date

plots = list()
for (i in 1:length(country_names)) {
  plots[[country_names[i]]] = hml.sma.5[, c(i, 7)] %>% 
    setNames(c('ret', 'date')) %>%
    ggplot() + 
    geom_bar(aes(x = as.factor(date), y = ret), stat = 'identity') + 
    theme(axis.text.x = element_text(angle = 90)) +
    ylab("HML Returns SMA5") +
    ggtitle(country_names[i])
}

do.call("grid.arrange", c(plots, ncol = 2))
```

```{r}
p1 = hml.sma.5 %>% 
  ggplot(aes(x = as.factor(date))) +
  geom_bar(aes(y = aus, fill = "aus"), alpha = 0.5, stat = 'identity') +
  geom_bar(aes(y = fr, fill = 'fr'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ger, fill = 'ger'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ita, fill = 'ita'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = jpn, fill = 'jpn'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = uk, fill = 'uk'), alpha = 0.5, stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("HML Returns SMA5")

print(p1)

p2 = hml.sma.5 %>% filter(date > '200701' & date < '200901') %>% 
  ggplot(aes(x = as.factor(date))) + 
  geom_bar(aes(y = aus, fill = "aus"), alpha = 0.5, stat = 'identity') +
  geom_bar(aes(y = fr, fill = 'fr'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ger, fill = 'ger'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = ita, fill = 'ita'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = jpn, fill = 'jpn'), alpha = 0.5, stat = 'identity') + 
  geom_bar(aes(y = uk, fill = 'uk'), alpha = 0.5, stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("HML Returns SMA5") + 
  ggtitle("HML Returns SMA5 2007-2009")


print(p2)

```

Seems like value stocks performed better during this period.

# 3. Augment the international data with HML returns. Estimate the 7x7 correlation matrix of the value-growth spreads. Based on your results do you agree with the opinion of the casewriter of the DFA case?

```{r}
# Subset FF dates

hml$date = aus$date
merged = hml %>% merge(ff[, c('date', 'hml')], by.x = 'date', by.y = 'date') %>%
  rename(us = 'hml')

V1 = cor(merged[,-1])

V1 %>% kable() %>% kable_styling()

```

# 4. Given the correlations, would you recommend for US value investors to go global?

Yes, there seems to be low correlations and so we can improve our Sharpe ratio by diversifying internationally.


# 5. Combine the 7 value-growth spreads into a portfolio with minimum total risk. Compare the portfolio's standard deviation with that of HML. Any surprise?

```{r}
I = matrix(1, 7, 1)
# Weightings of Minimum Variance Portfolio
w1 = solve(V1) %*% I / sum(t(I) %*% solve(V1))

t(w1) %>% kable() %>% kable_styling()
```

Very low weighting of 6.5% on US equities. Seems like international stocks offer lower variances.

# 6. Consider a mean-variance investor who currently holds all her $1 million in the US market portfolio. She holds zero position in the risk-free T-bill and she considers such a position optimal, given the historical moments of asset returns. What is the investor's risk aversion $\gamma$? Explain how you obtained your result.

```{r}
# Filter after 1975
ff = ff %>% filter(date >= '197501')
V = var(ff$mkt / 100)
E = mean(ff$mkt / 100)

gamma = E/V


print(gamma)


```
$w_{op} = \frac{1}{\gamma} V^{-1} E$ Need to solve for $\gamma$

# 7. This investor is considering tilting her portfolio toward value stocks. She is unwilling to invest internationally, but she is willing to invest in HML. Using sample estimates of expected returns, variances, and covariances, compute the optimal allocation of this investor across the US market, HML, and the T-bill.

```{r}
assets = ff[,c("mkt", "hml")]

E = as.matrix(colMeans(assets))
V = as.matrix(cov(assets))

W = 1 / gamma * 100 * solve(V) %*% E
t(W) %>% kable() %>% kable_styling()

```

Short the T-bill to buy more stocks


# 8. How much did the investor's Sharpe ratio increase as a result of the value tilt? How much did the squared Sharpe ratio increase?

```{r}
sr = mean(ff$mkt) / sd(ff$mkt)

print(paste("Market SR:", sr))


sr2 = t(W) %*% E / sqrt(t(W) %*% V %*% W)

print(paste("HML SR:", sr2[1]))

print(paste("SR diff: ", sr2[1] - sr))

print(paste("SR sq diff:", sr2[1]**2 - sr**2))

```

# 9. Compute HML's information ratio relative to the US market. How does the square of this ratio compare to the improvement in the squared Sharpe ratio that you computed in the previous part?

```{r}

inf.lm = lm(hml ~ mkt, data = ff)
alpha = inf.lm$coefficients[1]
se = sd(inf.lm$residuals)

IR = alpha / se

print(paste("Information ratio: ", IR))

print(paste("Information ratio sq: ", IR**2))




```
Same as the Sharpe ratio squared diff.


# 10. Now suppose this investor is willing to take positions in any of the seven international value-growth spreads. Using the sample estimate as before, construct the investor's optimal portfolio of the eight risky assets.

```{r}

assets = cbind(ff[, c("mkt", 'hml')], hml[, 1:6]) %>% rename(us = 'hml')

E = colMeans(assets) %>% as.matrix()

V = cov(assets) %>% as.matrix()


W = 1 / gamma * 100 * solve(V) %*% E 

t(W) %>% kable() %>% kable_styling()

```

# 11. Compare the resulting Sharpe ratio with the Sharpe ratio obtained by the US only optimal combination of the market with HML. Given this comparison, would you recommend for US value investors to go global?

```{r}

sr3 = t(W) %*% E / sqrt(t(W) %*% V %*% W)

print(paste("Sharpe ratio: ", sr3[1]))

```

Yes, definitely go global.





