---
title: "Homework 6"
author: "Tony Lian"
output: html_document
---
  
```{r setup, include=FALSE}
options(scipen = 10)
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
library(moments)
library(lubridate)
library(gridExtra)
library(TTR)


liq = read.table("data/liq_data_1962_2018.txt", skip = 11, header= FALSE)
names(liq) = c("date", "agg_liq", "innov_liq", "traded_liq")

liqvw = read.table("data/liq_vw_hist_deciles_1968_2018.txt", skip = 2, header = FALSE)
names(liqvw) = c("d1", "d2", "d3", "d4", "d5", "d6", "d7", "d8", "d9", "d10", "d10m1")

liqvw$index = c(1:nrow(liqvw)) - 1
liqvw$month = liqvw$index %% 12 + 1
liqvw$year = 1968 + floor(liqvw$index / 12)
liqvw$date = liqvw$year * 100 + liqvw$month

ff = read.table("data/ff_factors_192607_201912.txt", header = TRUE)
names(ff) = c("date", "mkt", "smb", "hml", "rf")

vix = read.table("data/vix.txt", skip=1)
names(vix) = c("vix")

vix$index = c(1:nrow(vix)) -1
vix$month = (vix$index) %% 12 + 1
vix$year = 1990 + floor(vix$index/12)
vix$date = vix$year * 100 + vix$month

```

# 1. What is the correlation between the innovations in liquidity and the VIX index? Use the longest time period over which both series are available. Provide an economic explanation for why this correlation is negative.

```{r}
m = merge(vix, liq, by = 'date')

c = cor(m$vix, m$innov_liq)

print(paste("Correlation:", c))

```
Increased innovations in liquidity allow for less friction in markets and leads to higher efficiency which also leads to decreased volatility as bid/ask spreads are tighter.


# 2. What is the correlation between the innovations in liquidity and stock market returns? Can you explain the sign of the correlation?

```{r}
m2 = merge(ff, liq, by = 'date')

c2 = cor(m2$mkt, m2$innov_liq)

print(paste("Correlation:", c2))

```

## a) Calculate the same correlation in two distinct subsamples: downmarkets and upmarkets.

```{r}

dn = m2[m2$mkt < 0, ]
up = m2[m2$mkt > 0, ]

dn_c = cor(dn$mkt, dn$innov_liq)
up_c = cor(up$mkt, up$innov_liq)

print(paste("Down cor:", dn_c))
print(paste("Up cor:", up_c))

```
There is a difference it seems like. 



# 3. Is liquidity risk priced? Do stocks with higher liquidity betas continue to have higher Fama-French alphas? Does the 10-1 porfolio have a significantly positive alpha?

```{r}
# calculate excess 

port = merge(ff, liqvw, on = date)
port = merge(port, liq, on = date)

for (i in 1:10) {
  port[paste0("exd", i)] = port[paste0("d", i)] * 100 - port$rf
}

port$d10m1 = port$d10m1 * 100

port.all = port[port$date > 196200, ]


reg_table = as.data.frame(matrix(NA, 2, 11,
                                 dimnames = list(c('alpha', 't'),
                                                 names(liqvw)[1:11])))


for (i in 1:10) {
  
  reg = lm(paste0("exd", i, " ~ mkt + smb + hml"), data = port.all)
  
  # Store alpha annualized
  reg_table[1, i] = reg$coefficients[1] * 12
  
  # Store beta for liquidity
  reg_table[2, i] = summary(reg)$coefficients[1, 3]
 
}

# For d10m1
reg = lm("d10m1 ~ mkt + smb + hml", data = port.all)
reg_table[1, 11] = reg$coefficients[1] * 12
reg_table[2, 11] = summary(reg)$coefficients[1, 3]

reg_table %>% kable() %>% kable_styling()


```

## a) Answer the same questions using the sample period Jan 2000 - Dec 2018. 

```{r}
port.oos = port[port$date > 200000,]

reg_table_oos = as.data.frame(matrix(NA, 2, 11,
                                 dimnames = list(c('alpha', 't'),
                                                 names(liqvw)[1:11])))


for (i in 1:10) {
  
  reg = lm(paste0("exd", i, " ~ mkt + smb + hml"), data = port.oos)
  
  # Store alpha annualized
  reg_table_oos[1, i] = reg$coefficients[1] * 12
  
  # Store beta for liquidity
  reg_table_oos[2, i] = summary(reg)$coefficients[1, 3]
 
}

# For d10m1
reg = lm("d10m1 ~ mkt + smb + hml", data = port.oos)
reg_table_oos[1, 11] = reg$coefficients[1] * 12
reg_table_oos[2, 11] = summary(reg)$coefficients[1, 3]

reg_table_oos %>% kable() %>% kable_styling()


```


## b) For both time periods, plot the Fama-French alphas across the 10 portfolios.

```{r}
t_table = as.data.frame(t(reg_table))
t_table$portfolio = rownames(t_table)

t_table %>% 
  ggplot(aes(x = portfolio, y = alpha)) + geom_bar(stat = 'identity')


```


# 4. Do historical liquidity betas predict future liquidity betas?  For both periods, plot those liquidity betas across the 10 portfolios. Do the betas exhibit an increasing pattern across the 10 portfolios? Does the 10-1 portfolio have a significantly positive liquidity beta?

```{r}

# Betas

liq_table_all = as.data.frame(matrix(NA, 2, 11,
                                     dimnames = list(c('beta', 't'), 
                                                     names(liqvw)[1:11])))


for (i in 1:10) {
  
  reg = lm(paste0("exd", i, " ~ innov_liq + mkt + smb + hml"), data = port.all)
  
  # Store alpha annualized
  liq_table_all[1, i] = reg$coefficients[2]
  
  # Store beta for liquidity
  liq_table_all[2, i] = summary(reg)$coefficients[2, 3]
 
}

# For d10m1
reg = lm("d10m1 ~ innov_liq + mkt + smb + hml", data = port.all)
liq_table_all[1, 11] = reg$coefficients[1] * 12
liq_table_all[2, 11] = summary(reg)$coefficients[2, 3]


liq_table_all %>% kable() %>% kable_styling()


# Out of Sample

liq_table_oos = as.data.frame(matrix(NA, 2, 11,
                                     dimnames = list(c('beta', 't'), 
                                                     names(liqvw)[1:11])))


for (i in 1:10) {
  
  reg = lm(paste0("exd", i, " ~ innov_liq + mkt + smb + hml"), data = port.oos)
  
  # Store alpha annualized
  liq_table_oos[1, i] = reg$coefficients[2]
  
  # Store beta for liquidity
  liq_table_oos[2, i] = summary(reg)$coefficients[2, 3]
 
}

# For d10m1
reg = lm("d10m1 ~ innov_liq + mkt + smb + hml", data = port.oos)
liq_table_oos[1, 11] = reg$coefficients[1] * 12
liq_table_oos[2, 11] = summary(reg)$coefficients[2, 3]



liq_table_oos %>% kable() %>% kable_styling()

```

```{r}

plot_beta_all = liq_table_all %>% t() %>% as.data.frame()
plot_beta_all$portfolio = rownames(plot_beta_all)


plot_beta_all %>% 
  ggplot(aes(x = portfolio, y = beta)) + geom_bar(stat = 'identity')


plot_beta_oos = liq_table_oos %>% t() %>% as.data.frame()
plot_beta_oos$portfolio = rownames(plot_beta_oos)

plot_beta_oos %>%
  ggplot(aes(x = portfolio, y = beta)) + geom_bar(stat = 'identity')


```


